Tesseroids: forward modeling gravitational fields in spherical coordinates

### 2015-03-30 manuscript.tex
White edges in split figure
Comments by Valeria and Carla

Satellite missions dedicated to measuring the Earth's gravity field
(like CHAMP, GRACE, and GOCE)
have provided geophysicists with almost uniform and global data coverage.
These new data have enabled interpretations on regional and global scales
\citep[e.g.][]{Reguzzoni2013,Braitenberg2015}.
Modeling at such scales requires taking into account the curvature of the
Earth and calculating gravity gradients as well as the traditional
gravitational acceleration.
A common approach to achieve this is
to discretize the Earth into tesseroids (Figure~\ref{fig:tesseroid})
instead of rectangular prisms.
However, the integral formula for the gravitational effects of a tesseroid
must be solved numerically.
Approaches to this numerical integration include
Taylor series expansion \citep{Heck2007, Grombein2013}
and the Gauss-Legendre Quadrature \citep{Asgharzadeh2007}.
Taylor series expansion produces accurate results at low latitudes but
presents a decrease in accuracy towards the polar regions.
This is attributed to tesseroids degenerating into an approximately triangular
shape at the poles.
The Gauss-Legendre Quadrature (GLQ) integration consists of approximating
the volume integral by a weighted sum of the effect of point masses.
An advantage of the GLQ approach is that it can be
controlled by the number of point masses used.
The larger the number of point masses,
the better the accuracy of GLQ integration.
A disadvantage is the increased computation time as the number of
point masses increases.
Thus, there is a trade-off between accuracy and computation time.
This is a common theme in numerical methods.
\citet{Wild-Pfeiffer2008} investigated the use of different mass elements,
including tesseroids, to compute the gravitational effects of topographic
masses.
The author concludes that using tesseroids with GLQ integration gives the best
results for near-zone computations.
However, the question of how to determine the optimal parameters for GLQ
integration remained open.

Previous work by \citet{Ku1977} investigated the use of the GLQ
in gravity forward modeling.
\citet{Ku1977} numerically integrated the vertical component of the
gravitational acceleration of right rectangular prisms.
The author suggested that the accuracy of the GLQ integration depends on
the ratio between distance to the computation point and the distance between
adjacent point masses.
Based on this, \citet{Ku1977} proposed an empirical criterion that the distance
% * <berg@units.it> 2015-03-19T13:46:11.161Z:
%
%  an criterion or proposed empirical criteria
%
% ^ <leouieda@gmail.com> 2015-03-20T17:41:12.218Z:
%
%  It was an criterion
%
between adjacent point masses should be less than
the distance to the computation point.
\citet{Asgharzadeh2007} used  this criteria for the GLQ integration of
the gravity gradient tensor of tesseroids.
To our knowledge, an analysis of how well this ad hoc criteria of
\citet{Ku1977} works for gravity gradient components or for tesseroids
has never been done before.
There has also been no attempt to quantify the error committed in the GLQ
integration when applying the criteria of \citet{Ku1977}.

\citet{Li2011} devised an algorithm to automatically enforce the criteria of
\citet{Ku1977}.
Their algorithm divides the tesseroid into smaller ones instead of increasing
the number of point masses per tesseroid.
% * <berg@units.it> 2015-03-19T13:48:33.551Z:
%
%  sentence is repeated later.
%
% ^ <leouieda@gmail.com> 2015-03-20T17:40:17.120Z:
%
%  I removed it.
%
A tesseroid is divided if the minimum distance to the computation point
is smaller than the largest dimension of the tesseroid.
This division is repeated recursively until all tesseroids obey the criterion.
Then, GLQ integration is performed for each of the smaller tesseroids
using the specified number of point masses.
The advantage of this adaptive discretization over
increasing the number of points masses is that the
total distribution of point masses will be greater
only close to the computation point.
This makes the adaptive discretization more computationally efficient.

\citet{Grombein2013} developed optimized formula for the gravitational fields
of tesseroids using Cartesian integral kernels.
These formulas are faster to compute and do not have singularities at the poles
like their spherical counterparts.
The Cartesian formulae are numerically integrated using a Taylor series
% * <berg@units.it> 2015-03-19T13:49:40.239Z:
%
%  singular or plural?
%
% ^ <leouieda@gmail.com> 2015-03-20T17:41:54.790Z:
%
%  It should be the plural (formulae)
%
expansion as per \citet{Heck2007}.
\citet{Grombein2013} use a near-zone separation to mitigate the increased error
at high latitudes.
In the so called ``near-zone'' of the computation point they use a finer
discretization composed by smaller tesseroids.
This is accomplished by dividing the tesseroids along their horizontal
dimensions.
However, the determination of an optimal size of the near-zone remains an
open question \citep{Grombein2013}.

@@

This article describes the software design and the implementation
of our modified adaptive discretization algorithm.
We also present a numerical investigation of the error committed
in the computations.
These results allow us calibrate the adaptive discretization algorithm
separately for the gravitational potential, gravitational acceleration,
as well as the gravity gradient tensor components.

@@

\noindent
where $\alpha, \beta \in \{x, y, z\}$,
$\rho$ is the density,
$G = 6.673\times10^{-11}\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
and

@@

We will follow \citet{Asgharzadeh2007} and perform the numerical integration
using the Gauss-Legendre Quadrature (GLQ).
The GLQ consists of approximating the integral by a weighted sum of the
integration kernel \citep{Hildebrand1987},

@@

Our implementation of the adaptive discretization algorithm
differs in a few ways from one proposed by \citet{Li2011}.
In \citet{Li2011},
a tesseroid will be divided when
the smallest distance between it and the computation point
is smaller than a constant times
the largest dimension of the tesseroid.
Instead of the smallest distance,
we use the easier to calculate
distance between
the computation point $(r, \lambda, \phi)$
and the geometric center of the tesseroid
$(r_t, \lambda_t, \phi_t)$

@@

\begin{enumerate}
    \item ``Pop'' the stack (i.e., take the last tesseroid from it).
        This will cause \emph{stktop} to be reduced by one.
        This tesseroid is the one that will be evaluated in the following
        steps.
    \item Compute the distance $d$ (equation~\ref{eq:distance}) between
        the geometric center of the tesseroid and
        the computation point.
    \item Compute the dimensions of the tesseroid $L_\lambda$, $L_\phi$,
        and $L_r$ using equations~\ref{eq:sizelon}-\ref{eq:sizer}.
    \item Check the condition in equation~\ref{eq:condition} for each
        dimension of the tesseroid.
    \item If all dimensions hold the inequality \ref{eq:condition},
        the tesseroid is not divided and its
        gravitational effect is computed using the
        Gauss-Legendre Quadrature
        (equations~\ref{eq:tesspot}-\ref{eq:tesstensor} and~\ref{eq:glq3d}).
        We use a GLQ order of two for all three dimensions
        ($2 \times 2 \times 2 = 8$ point masses)
        by default.
        This value can be changed using
        a command-line argument of the modeling programs.
    \item If any of the dimensions fail the condition:
    \begin{enumerate}
        \item Divide the tesseroid in half along each dimension that failed
             the condition.
        \item Check if there is room in the stack
            for the new tesseroids
            (i.e.,the number of new elements plus \emph{stktop}
             is smaller than the maximum stack size).
             If there isn't,
             warn the user of a ``stack overflow''
             and compute the effect of the larger tesseroid (step 5).
             If there is room in the stack,
             place the smaller tesseroids into
             the stack.
    \end{enumerate}
    \item Repeat the above steps until the stack is empty
        (\emph{stktop} is equal to -1).
\end{enumerate}

The algorithm above is repeated
for every tesseroid of the input model
and the results are summed.
This will yield the gravitational effect
of the input tesseroid model on a single point.
Thus, the computations must be repeated
for every computation point.
The whole algorithm can be summarized
in the following pseudo-code.

@@

Figure~\ref{fig:glqerrorsample} shows that the largest errors are spread over
on top of the tesseroid.
Thus, calculating the tesseroid fields at a single point might not
capture the point of largest error.
Instead, we calculate the effect of the tesseroid model on a regular grid
of $10 \times 10$ points at different geographic locations
(see Table~\ref{tbl:experiment}).
Fortunately, the symmetry of the shell allows us to consider the computation
point at any geocentric coordinate.
Therefore, the effect of the shell will be same along the entire grid.
We compute the differences between the effects of the shell and the tesseroid
model on the grid.
However, we will consider only the largest error in our analysis.

We placed the grid on top of a particular tesseroid
to increase the chances of capturing the true largest integration error.
We calculate the errors for values of the distance-size ratio
$D$ varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
Furthermore, we repeated the error analysis in four different numerical experiments,
each with computation grids at different locations and
different tesseroid model sizes.
Table~\ref{tbl:experiment} describes the different numerical
experiments and the corresponding parameters of the computation grid and
tesseroid model.

Figure~\ref{fig:dist-size-curves} shows
the maximum difference between the shell and tesseroid fields
as a function of $D$ for the four experiments.
The differences are given as a percentage of the shell value.
We established a maximum tolerated error of $0.1\%$, represented by the
horizontal solid lines in Figure~\ref{fig:dist-size-curves}.
Only results for the gravitational potential, $g_z$, and $g_{zz}$ are shown.
The results for the other diagonal components of the gravity gradient tensor
are similar to $g_{zz}$.
Figures for these components can be found in the supplementary material
(see section "Reproducing the analysis and results").

For the potential $V$, a distance-size ratio $D=1$ guarantees that the curves
for all experiments are below the $0.1\%$ error threshold.
For $g_z$, the same is achieved with $D=1.5$.
Conversely, $g_{zz}$ requires a value of $D=8$ to achieve an error level of
$0.1\%$.
For a computation height of 260 km, the error curve for $g_{zz}$
intercepts the error threshold line at $D=2.5$.
This behavior suggests that the error curves for $g_{zz}$ might depend on the
computation height.
To test this hypothesis, we computed the error curves for $g_{zz}$ at
heights 2, 10, 50, 150, and 260 km.
Figure~\ref{fig:gzz-with-height} shows the results for $g_{zz}$ at varying
computation heights.
Notice that the distance-size ratio required to achieve $0.1\%$ accuracy
decreases as the computation height increases.
% * <berg@units.it> 2015-03-19T21:07:33.445Z:
%
%  what is the height dependence due to? Should the distance not include height already? Is the height a second control parameter, next to the distance ratio? What is that due to? Is it only for gzz or also for the other gradient components? What about the gx?
%
% ^ <leouieda@gmail.com> 2015-03-20T18:23:53.901Z:
%
%  The height is included in the distance and is not a control variable. This happens for all tensor components but not to the potential, gx, gy, and gz. I think this is because the decay factor for the tensor components is 1/r^3 whereas the discreatization algorithm compares D to r/size. So as the  computation point becomes closer. the field increases more rapidly than the algorithm increases the  amount of discretization. That is why you need higher D (more discretization) when you are closer to the tesseroid.
%
For example, computation at 260 km height requires $D=2.5$ whereas at 10 km
height a value of $D=5.5$ is required to achieve the same accuracy.
One can take advantage of this behavior to reduce the distance-size ratio for
computations of the gravity gradient tensor at high altitudes,
saving computation time.

>>>>

Satellite missions dedicated to measuring the Earth's gravity field
(like CHAMP, GRACE, and GOCE)
have provided geophysicists with almost uniform and global data coverage.
These new data have enabled interpretations on regional and global scales
\citep[e.g.][]{Reguzzoni2013,Braitenberg2015}.
Modeling at such scales requires taking into account the curvature of the
Earth and calculating gravity gradients as well as the traditional
gravitational acceleration.
A common approach to achieve this is
to discretize the Earth into tesseroids (Figure~\ref{fig:tesseroid})
instead of rectangular prisms.
However, the integral formula for the gravitational effects of a tesseroid
must be solved numerically.
Approaches to this numerical integration include
Taylor series expansion \citep{Heck2007, Grombein2013}
and the Gauss-Legendre Quadrature \citep{Asgharzadeh2007}.
Taylor series expansion produces accurate results at low latitudes but
presents a decrease in accuracy towards the polar regions.
This is attributed to tesseroids degenerating into an approximately triangular
shape at the poles.
The Gauss-Legendre Quadrature (GLQ) integration consists in approximating
the volume integral by a weighted sum of the effect of point masses.
An advantage of the GLQ approach is that it can be
controlled by the number of point masses used.
The larger the number of point masses,
the better the accuracy of GLQ integration.
A disadvantage is the increased computation time as the number of
point masses increases.
Thus, there is a trade-off between accuracy and computation time.
This is a common theme in numerical methods.
\citet{Wild-Pfeiffer2008} investigated the use of different mass elements,
including tesseroids, to compute the gravitational effects of topographic
masses.
The author concludes that using tesseroids with GLQ integration gives the best
results for near-zone computations.
However, the question of how to determine the optimal parameters for GLQ
integration remained open.

Previous work by \citet{Ku1977} investigated the use of the GLQ
in gravity forward modeling.
\citet{Ku1977} numerically integrated the vertical component of the
gravitational acceleration of right rectangular prisms.
The author suggested that the accuracy of the GLQ integration depends on
the ratio between distance to the computation point and the distance between
% * <valeria.valcris@gmail.com> 2015-03-23T17:28:28.769Z:
%
%  Leo, the sentence "Based on this, \citet{Ku1977} proposed an empirical criterion that the distance the distance to the computation point."
%
% ^ <valeria.valcris@gmail.com> 2015-03-23T17:57:28.416Z.
adjacent point masses.
% * <valeria.valcris@gmail.com> 2015-03-23T17:29:37.800Z:
%
%  is wrong. The criterion disappeared.
%
% ^ <valeria.valcris@gmail.com> 2015-03-23T17:57:27.222Z.
Based on this, \citet{Ku1977} proposed an empirical criterion that
the distance between point masses should be greater than
the distance to the computation point.
\citet{Asgharzadeh2007} used  this criterion for the GLQ integration of
the gravity gradient tensor of tesseroids.
To our knowledge, an analysis of how well this ad hoc criteria of
\citet{Ku1977} works for gravity gradient components or for tesseroids
has never been done before.
There has also been no attempt to quantify the error committed in the GLQ
integration when applying the criteria of \citet{Ku1977}.

\citet{Li2011} devised an algorithm to automatically enforce the criteria of
\citet{Ku1977}.
Their algorithm divides the tesseroid into smaller ones instead of increasing
the number of point masses per tesseroid.
A tesseroid is divided if the minimum distance to the computation point
is smaller than the largest dimension of the tesseroid.
This division is repeated recursively until all tesseroids obey the criterion.
Then, GLQ integration is performed for each of the smaller tesseroids
using the specified number of point masses.
The advantage of this adaptive discretization over
increasing the number of points masses is that the
total distribution of point masses will be greater
only close to the computation point.
This makes the adaptive discretization more computationally efficient.

\citet{Grombein2013} developed optimized formula for the gravitational fields
of tesseroids using Cartesian integral kernels.
These formulas are faster to compute and do not have singularities at the poles
like their spherical counterparts.
The Cartesian formulae are numerically integrated using a Taylor series
expansion as per \citet{Heck2007}.
\citet{Grombein2013} use a near-zone separation to mitigate the increased error
at high latitudes.
In the so called ``near-zone'' of the computation point they use a finer
discretization composed by smaller tesseroids.
This is accomplished by dividing the tesseroids along their horizontal
dimensions.
However, the determination of an optimal size of the near-zone remains an
open question \citep{Grombein2013}.

@@

This article describes the software design and the implementation
of our modified adaptive discretization algorithm.
We also present a numerical investigation of the error committed
in the computations.
These results allow us to calibrate the adaptive discretization algorithm
separately for the gravitational potential, gravitational acceleration,
as well as the gravity gradient tensor components.

@@

\noindent
where $\alpha, \beta \in \{x, y, z\}$,
$\rho$ is the density,
$G = 6.674\times10^{-11}\ m^3kg^{-1}s^{-1}$ is the gravitational constant,
% * <berg@units.it> 2015-03-22T10:56:54.365Z:
%
%  Leo, on the link below the digits are 6.67384, so if you truncate it should be 6.674 not 6.673 as on the wikipedia article on G.
%
% * <berg@units.it> 2015-03-22T10:54:00.463Z:
%
%  http://physics.nist.gov/cgi-bin/cuu/Value?bg|search_for=universal_in!
%
% ^ <leouieda@gmail.com> 2015-03-22T21:20:28.891Z:
%
%  That's correct, thanks. It must have been a typing error.
%
and

@@

We will follow \citet{Asgharzadeh2007} and perform the numerical integration
using the Gauss-Legendre Quadrature (GLQ).
The GLQ consists in approximating the integral by a weighted sum of the
integration kernel \citep{Hildebrand1987},

@@

Our implementation of the adaptive discretization algorithm
differs in a few ways from the one proposed by \citet{Li2011}.
In \citet{Li2011},
a tesseroid will be divided when
the smallest distance between it and the computation point
is smaller than a constant times
the largest dimension of the tesseroid.
Instead of the smallest distance,
we use the easier to calculate
distance between
the computation point $(r, \lambda, \phi)$
and the geometric center of the tesseroid
$(r_t, \lambda_t, \phi_t)$

@@

\begin{enumerate}
    \item ``Pop'' the stack (i.e., take the last tesseroid from it).
        This will cause \emph{stktop} to be reduced by one.
        This tesseroid is the one that will be evaluated in the following
        steps.
    \item Compute the distance $d$ (equation~\ref{eq:distance}) between
        the geometric center of the tesseroid and
        the computation point.
    \item Compute the dimensions of the tesseroid $L_\lambda$, $L_\phi$,
        and $L_r$ using equations~\ref{eq:sizelon}-\ref{eq:sizer}.
    \item Check the condition in equation~\ref{eq:condition} for each
        dimension of the tesseroid.
    \item If all dimensions hold the inequality \ref{eq:condition},
        the tesseroid is not divided and its
        gravitational effect is computed using the
        Gauss-Legendre Quadrature
        (equations~\ref{eq:tesspot}-\ref{eq:tesstensor} and~\ref{eq:glq3d}).
        We use a GLQ order of two for all three dimensions
        ($2 \times 2 \times 2 = 8$ point masses)
        by default.
        This value can be changed using
        a command-line argument of the modeling programs.
    \item If any of the dimensions fail the condition:
    \begin{enumerate}
        \item Divide the tesseroid in half along each dimension that failed
             the condition.
        \item Check if there is room in the stack
            for the new tesseroids
            (i.e.,the number of new elements plus \emph{stktop}
             is smaller than the maximum stack size).
             If there isn't,
             warn the user of a ``stack overflow''
             and compute the effect of the larger tesseroid (step 5).
             If there is room in the stack,
             place the smaller tesseroids into
             the stack.
    \end{enumerate}
    \item Repeat the above steps until the stack is empty
        (\emph{stktop} is equal to -1).
\end{enumerate}
% * <valeria.valcris@gmail.com> 2015-03-23T19:53:50.437Z:
%
%  Leo, item 6b I suggest changing " warn the user of a ``stack overflow''              and compute the effect of the larger tesseroid (step 5)." by
%
The algorithm above is repeated
% * <valeria.valcris@gmail.com> 2015-03-23T21:09:58.768Z:
%
%  by "warn the user of a ``stack overflow''  and compute the effect of the tesseroid computed on the step 5."
%
for every tesseroid of the input model
and the results are summed.
This will yield the gravitational effect
of the input tesseroid model on a single point.
Thus, the computations must be repeated
for every computation point.
The whole algorithm can be summarized
in the following pseudo-code.

@@

Figure~\ref{fig:glqerrorsample} shows that the largest errors are spread over
on top of the tesseroid.
Thus, calculating the tesseroid fields at a single point might not
capture the point of largest error.
Instead, we calculate the effect of the tesseroid model on a regular grid
of $10 \times 10$ points at different geographic locations
% * <berg@units.it> 2015-03-22T12:57:56.006Z:
%
%  and takethe average over the hundert points. Leo: is that correct? Should we add it to be more clear?
%
% ^ <leouieda@gmail.com> 2015-03-22T20:37:47.094Z:
%
%  No, I don't take the average. I take the largest error. I didn't mention that above because I only talk about the error further on ("Figure~\ref{fig:dist-size-curves} shows the maximum difference between the shell and tesseroid fields").
%
(see Table~\ref{tbl:experiment}).
Fortunately, the symmetry of the shell allows us to consider the computation
point at any geocentric coordinate.
Therefore, the effect of the shell will be same along the entire grid.
We compute the differences between the effects of the shell and the tesseroid
model on the grid.
However, we will consider only the largest error in our analysis.

We placed the grid on top of a particular tesseroid
to increase the chances of capturing the true largest integration error.
We calculate the errors for values of the distance-size ratio
$D$ varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.
% * <berg@units.it> 2015-03-22T13:00:05.803Z:
%
%  D=0 means we are sitting in the center of the tesseroid, more than no division. Maybe you could be more precise on this, as D=0 will not happen, as you are at 2km height?
%
% ^ <leouieda@gmail.com> 2015-03-22T20:45:02.247Z:
%
%  I'm not moving the computation grid during this test. What I do is I vary the value of D used in the adaptive discretization. That's what is shown in Fig 4. Each curve corresponds to a fixed grid. Changing D just means that the adaptive discretization is performed differently because of the inequality in eq 17. Notice that if D=0, any combination of distance/size is larger than D. So there won't be any divisions.
%
Furthermore, we repeated the error analysis in four different numerical experiments,
each with computation grids at different locations and
different tesseroid model sizes.
Table~\ref{tbl:experiment} describes the different numerical
experiments and the corresponding parameters of the computation grid and
tesseroid model.

Figure~\ref{fig:dist-size-curves} shows
the maximum difference between the shell and tesseroid fields
as a function of $D$ for the four experiments.
The differences are given as a percentage of the shell value.
We established a maximum tolerated error of $0.1\%$, represented by the
horizontal solid lines in Figure~\ref{fig:dist-size-curves}.
Only results for the gravitational potential, $g_z$, and $g_{zz}$ are shown.
% * <berg@units.it> 2015-03-22T13:02:41.484Z:
%
%  Not sure I understand what you did for the 30° tesseroid. Errors are nearly the same. It means that at D=1 the tesseroid has been divided into many small tesseroids so as to have the correct distance-size ratio: Already for D=.5 the calculation is made on small tesseroids that make up the big 30° tesseroid. So actually the different curves represent very similar  tesseroids in the near field, only in the far field are the tesseroids left with the large size. We see that the result is nearly the same, with a great reduction in computer time in the case of the 30° tesseroid as for distant tesseroids no division is necessary. Leo: maybe we could add this comment, if my interpretation is correct.
%
% ^ <leouieda@gmail.com> 2015-03-22T21:19:24.057Z:
%
%  Carla, your interpretation is correct. In practice there is no advantage to using large tesseroids with this algorithm because I don't take into account a finer resolution model in the near zone (like a proper near-far zone separation).  The difference is that I'm dividing a corse model into small but equivalent tesseroids. For a near-far zone separation I'd average out the fine model in the far zone.
%
The results for the other diagonal components of the gravity gradient tensor
are similar to $g_{zz}$.
Figures for these components can be found in the supplementary material
(see section "Reproducing the analysis and results").

For the potential $V$, a distance-size ratio $D=1$ guarantees that the curves
for all experiments are below the $0.1\%$ error threshold.
For $g_z$, the same is achieved with $D=1.5$.
Conversely, $g_{zz}$ requires a value of $D=8$ to achieve an error level of
$0.1\%$.
For a computation height of 260 km, the error curve for $g_{zz}$
intercepts the error threshold line at $D=2.5$.
This behavior suggests that the error curves for $g_{zz}$ might depend on the
computation height.
To test this hypothesis, we computed the error curves for $g_{zz}$ at
heights 2, 10, 50, 150, and 260 km.
Figure~\ref{fig:gzz-with-height} shows the results for $g_{zz}$ at varying
computation heights.
Notice that the distance-size ratio required to achieve $0.1\%$ accuracy
decreases as the computation height increases.
% * <berg@units.it> 2015-03-19T21:07:33.445Z:
%
%  what is the height dependence due to? Should the distance not include height already? Is the height a second control parameter, next to the distance ratio? What is that due to? Is it only for gzz or also for the other gradient components? What about the gx?
%
% ^ <leouieda@gmail.com> 2015-03-20T18:23:53.901Z:
%
%  The height is included in the distance and is not a control variable. This happens for all tensor components but not to the potential, gx, gy, and gz. I think this is because the decay factor for the tensor components is 1/r^3 whereas the discreatization algorithm compares D to r/size. So as the  computation point becomes closer. the field increases more rapidly than the algorithm increases the  amount of discretization. That is why you need higher D (more discretization) when you are closer to the tesseroid.
%
% ^ <berg@units.it> 2015-03-22T13:32:27.725Z:
%
%  hello Leo: are you sure D<2 is what he is calculating at 260km height? if it is a 1° tesseroid and you are at 260km height, your distance is greater than the biggest sidelength of the tesseroid, and maybe that is why the error starts to decrease for D>2.5, because for smalle D your condition is satisfied by itself .
%
% ^ <leouieda@gmail.com> 2015-03-22T21:14:00.081Z:
%
%  That is exactly it. For 1º tesseroids at 260km, there are no divisions for D < 2 because the condition is already satisfied. That's why the error curve is flat.
%
For example, computation at 260 km height requires $D=2.5$ whereas at 10 km
height a value of $D=5.5$ is required to achieve the same accuracy.
One can take advantage of this behavior to reduce the distance-size ratio for
computations of the gravity gradient tensor at high altitudes,
saving computation time.
